Latitudinal gradients in biotic niche breadth vary across ecosystem types.

### 2018-08-08 manuscript/correction.tex
uploading a shortened correction

\begin{spacing}{1.0}
\section*{Abstract}

Several properties of food webs---the networks of feeding links between
species---are known to vary systematically with the species richness of the
underlying community.  Under the ``latitude-niche breadth hypothesis'', which
predicts that species in the tropics will tend to evolve narrower niches, one
might expect that these scaling relationships could also be affected by
latitude. To test this hypothesis, we analysed the scaling relationships
between species richness and average generality, vulnerability, and links per
species across a set of 196 empirical food webs. In estuarine, marine, and 
terrestrial food webs there was no effect of latitude on any scaling relationship,
suggesting constant niche breadth in these habitats. In freshwater communities,
on the other hand, there were strong effects of latitude on scaling relationships,
supporting the latitude-niche breadth hypothesis. These contrasting findings
indicate that it may be more important to account for habitat than latitude
when exploring gradients in food-web structure.

@@

% \linenumbers

\section*{Introduction}

  Food webs --networks of feeding links between species-- have been used for
  several decades to summarise the structure of  ecological
  communities~\cite{Paine1966,Williams2000,Petchey2008} and to understand how
  that structure relates to environmental variables  such as habitat
  type~\cite{Briand1983,Shurin2006}, primary
  productivity~\cite{Townsend1998,Thompson2005a,Vermaat2009}, and
  climate~\cite{Petchey2010,Baiser2012}. The latter variables in turn have
  strong gradients over latitude, with productivity and temperature both being
  higher in the tropics while  climate is more variable at high
  latitudes~\cite{Field1998}. These variables affect both the resources
  available and species'
  metabolisms~\cite{White2007,OConnor2009,Hechinger2011,White2011}, and have
  been proposed as determinants of the strength of  interspecific
  interactions~\cite{Schemske2009,Lang2012,Schleuning2012}.  By modulating
  interactions between species, latitudinal gradients may also shape food-web
  structure. Indeed, these latitudinal environmental gradients have been put
  forward as potential drivers for the  latitudinal gradient in species
  richness, one of the most general and robust patterns in
  ecology~\cite{Kaufman1995,Macpherson2002,Schemske2009}.

  One proposed link between species richness and latitude is the 
  ``latitude-niche breadth hypothesis''~\cite{Vazquez2004}.  This hypothesis 
  predicts that decreased seasonality in the tropics should lead to more stable
  populations, which in turn should evolve smaller niches~\cite{Vazquez2004}.
  These narrow niches should therefore allow more species to  coexist in the
  tropics than at higher latitudes. Alternatively, the higher productivity of
  the tropics~\cite{Brown2004} may result in a broader niche
  space~\cite{Davies2007} which could also sustain greater biodiversity even
  if niche  sizes are globally similar. Although the assumptions of the
  latitude-niche breadth hypothesis are only equivocally
  supported~\cite{Vazquez2004}, it remains a compelling potential mechanism
  for the latitudinal gradient in species
  richness~\cite{Lappalainen2006,Krasnov2008,Slove2010}.

  If the latitude-niche breadth hypothesis is correct, there should also be
  direct relationships between latitude and the degree of specialisation
  (i.e., the breadth of the Eltonian niche; ~\cite{Elton1927,Leibold2010}) of
  species within food webs. Specifically, narrower niches in the tropics would
  equate to greater specialisation (narrower niches) while constant niche
  sizes but greater productivity would translate to constant specialisation
  and niche width across latitude (Fig.~\ref{concept}). Attempts to unravel
  these effects, however, are complicated by known relationships between
  species richness and many other network properties~\cite{Riede2010}.  For
  example, narrower niches imply fewer links per species (i.e., greater
  specialisation) in the tropics (\cite{Marra1997,Dyer2007}; but
  see~\cite{Schleuning2012}). However, average numbers of links per species
  tend to increase in larger food webs~\cite{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialisation may be obscured by a latitudinal
  gradient in species richness.

  If this is the case, it may still be possible to uncover effects of latitude
  on specialisation by examining the shape of the scaling relationship between
  specialisation and species richness over changing latitude.  By testing
  whether latitude affects the scaling of each property with species richness,
  we test for the effects of latitude on specialisation predicted by the
  ``latitude-niche breadth hypothesis'' (Fig.~\ref{concept}).  If the scaling
  of specialisation with species richness is weaker in the  tropics (i.e., if
  species gain fewer links, predators, or prey as the size  of the network
  increases), this will indicate narrower niches at the tropics.  If, however,
  the scaling of specialisation with species richness does not  vary over
  latitude, this will indicate that niches are similarly-sized worldwide but
  that there is a broader niche space in the tropics. Additionally, as food
  webs describing different ecosystem types may differ in their
  topology~\cite{Dunne2004,Shurin2006}, we also explored the differences in
  scaling relationships across ecosystem types. Here, we use three measures of
  specialisation; mean links per species, mean generality (number of prey),
  and mean vulnerability (number of predators).

\section*{Methods}

  \subsection*{Data Set}

    We compiled a list of 196 empirical food webs from
    multiple sources (see \emph{Appendix S1} for web origins and selection
    criteria). We recorded study site latitude from the original source where
    possible or, where study sites were described but exact coordinates were not
    given, obtained estimated coordinates using Google Earth~\cite{GoogleEarth}.
    If a range of latitudes (e.g. $42-49^{\circ}N$) was provided, we used the midpoint
    of this range. We grouped food webs by ecosystem type (stream, N=71; lake, N=47; marine, 
    N=28; estuarine, N=18; and terrestrial, N=31) according to their designation in 
    previous aggregations of food webs (i.e.,~\cite{GlobalWeb,Riede2011,Dunne2013}).

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by using food webs comprised of ``trophic
    species'' --aggregations of species with identical sets of predators and prey-- rather than  
    species \emph{per se}~\cite{Martinez1991,Dunne2004,Vermaat2009,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original (i.e., without aggregating 
    any species) and trophic-species (i.e., after aggregating species with identical predators and prey) versions
    of the dataset; in each case using the number of species and 
    feeding links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness and specialisation across ecosystem type and latitude are very 
    similar whether or not species with identical sets of predators and prey are included. For simplicity, 
    here we present only the results for the original webs.

  \subsection*{Gradients over Latitude}

    To put our dataset in the context of other research on latitudinal gradients in species richness,
    we first examined simple linear relationships between latitude and each of 
    species richness, links per species, generality, and vulnerability. We fit models of the form

    \begin{equation}
    \label{Latfull}
    S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} ,
    \end{equation}

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks corresponding to $E_{i}=0$) and $\epsilon_{i}$ is a residual error term. 
    We next calculated the AIC
    of the maximal model as well as the AIC's of a suite of candidate simplified models identified
    using the R~\cite{R} function dredge from package MuMIn~\cite{MuMIn}. 
    Simplified models were obtained by
    systematically removing all possible combinations of terms from the full model.
    The best-fitting model was then determined to be the model with the fewest terms 
    where $\Delta$AIC\textless2, as this model is the least likely to suffer from over-fitting.

  \subsection*{Scaling Relationships with Species Richness}

    The scaling relationship between link density ($Z$) and species richness ($S$)
    has been shown to be a power law~\cite{Riede2010} of the form

    \begin{equation}
    \label{Power}
    Z_{i} \sim \alpha S_{i}^{\beta}  ,
    \end{equation}

    \noindent which is often re-expressed in logarithmic form

    \begin{equation}
    \label{Loglog}
    \log{Z_{i}} \sim \log{\alpha} + \beta\log{S_{i}}  .
    \end{equation}

    \noindent As the two forms imply a statistical fit of the data to different error 
    distributions, neither of which has strong
    \emph{a priori} justification in our dataset, we followed the recommendations in~\cite{Xiao2011}
    to compare the two model formulations explicitly (see \emph{Appendix S2} for details). 
    The logarithmic form (equation~\ref{Loglog}) provided the better fit to the data,
    as did the logarithmic forms of similar models for the scaling of generality and vulnerability. 
    We therefore used and present logarithmic models throughout the rest of the analyses.

  \subsection*{Effect of Latitude on Scaling}

    After determining the appropriate form of the scaling relationship, we
    then assessed the impact of latitude on the scaling relationships between
    species richness and link density, generality, vulnerability. In the
    context of the scaling relationships above, note  that this implies that
    we wished to determine the effect of latitude on the scaling exponent
    $\beta$. We included a categorical variable for
    ecosystem type (stream, lake,  terrestrial, marine, or estuary), as well
    as interactions between food web type and latitude.

    We therefore began by considering models of the form

    \begin{equation}
    \label{PowerLat}
    Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{2}E_{i}+\beta_{3}LE_{i}} + \epsilon_{i} ,
    \end{equation}

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks corresponding to $E_{i}=0$) and $\epsilon_{i}$ is a residual error term.
    The logarithmic formulation of this model is

    \begin{equation}
    \label{LogLat}
    \log{Z_{i}} = \log{\alpha}+\beta_{0}\log{S_{i}} + \beta_{1}L\log{S_{i}} +\beta_{2}E\log{S_{i}} +\beta_{3}LE\log{S_{i}} +\epsilon_{i} .
    \end{equation}

    We then simplified versions of model~\ref{LogLat} for link density, generality, and vulnerability following the procedure described above.
    As a supplemental check to ensure that variation in sampling effort
    across food webs was not responsible for the trends we observed, we then
    repeated our analyses using jackknifed data sets in which we 1) sequentially
    removed each food web in the dataset and 2) sequentially removed sets of food webs
    that shared a common author. The first jackknife essentially controls for the
    influence of any single outlier, while the second controls for the influence
    of particular research groups, some of which contributed large numbers of food
    webs (up to 27) to the dataset. Parameter estimates for the simplified models 
    varied very little across either jackknife test (see \emph{Appendix S3} for details),
    indicating that the trends we observed were not due to either strong outliers or
    to substantial differences in sampling effort across research groups.

\section*{Results}

  Link density (mean number of feeding links per species), generality (mean
  number of prey per species), and vulnerability (mean number of predators per
  species) were strongly and positively correlated ($R^2$=\st{0.891}0.915 for link
  density and generality, $R^2$\textgreater0.999 for link density and
  vulnerability, and $R^2$=\st{0.890}0.912 for generality and vulnerability). Contrary
  to the expected latitudinal gradient, the best-fit version of
  equation~(\ref{Latfull}) did not include a significant
  effect of latitude on species richness for any ecosystem type. Further, there were no significant
  relationships between link density, generality, or vulnerability with
  latitude for any ecosystem type.

  Each measure of specialisation increased with increasing
  species richness ($\beta_0$=\st{0.637}0.691, $p$\textless0.001; $\beta_0$=\st{0.553}0.663,
  $p$\textless0.001; and $\beta_0$=\st{0.637}0.698, $p$\textless0.001, respectively;
  Fig.~\ref{props_v_lat}). For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=\st{0.365}0.212 for link density;
  $\beta_{Latitude}$=-0.001, $p$=\st{0.535}0.320 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=\st{0.366}0.207 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=\st{0.004, $p$=0.019}0.003, $p$=0.037;
  $\beta_{Latitude:Lake}$=0.005, $p$=\st{0.004}0.007; and
  $\beta_{Latitude:Lake}$=\st{0.004, $p$=0.018}0.003, $p$=0.036, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=\st{0.007, $p$=0.001}0.005, $p$=0.008) while link density and
  vulnerability did not vary with latitude (i.e., the interaction term 
  $\beta_{Latitude:Stream}$ was not retained in the best-fit models).

\section*{Discussion}

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\cite{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009},
  intuitively one might suspect that any relationship between food-web
  properties such as generality might be due to the latitudinal gradient in
  species richness. In this dataset, however, we found no evidence to support
  latitudinal gradients in species richness, links per species, generality, 
  or vulnerability.

  The lack of a latitudinal gradient in species richness in this dataset
  contrasts strongly with other
  studies~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009}. As
  numbers of species and links included in a food web vary strongly with
  sampling effort as well as with the underlying diversity of the study area,
  it is possible that the lack of latitudinal trends here is a result of
  researchers tending to expend similar amounts of sampling effort across
  studies. This could result in food webs describing species-rich tropical
  communities omitting more species and links than studies of species-poor
  arctic communities if research groups spend similar person-hours assembling
  webs and can observe similar numbers of species and links per person-hour.
  In addition, it is worth noting that gradients in species richness are
  generally measured for a single taxonomic group at a
  time~\cite{Kaufman1995,Macpherson2002,Hillebrand2004,Schemske2009}. It is
  possible that these taxa are not well-represented in our food webs and that
  the dominant taxa that are represented do not have an underlying latitudinal
  gradient in richness. In either case, the lack of a strong association
  between observed species richness and latitude in any ecosystem type in our dataset
  means that any effect of latitude on other scaling relationships is not being driven by
  the scaling of specialisation with species richness in our food webs. This is fortunate since
  the lack of confounding effects of latitude allows us to more clearly assess
  effects of latitude on scaling with species richness.

  Scaling of links per species, generality, and vulnerability with species
  richness varied strongly across ecosystem types. In estuarine, marine, and
  terrestrial food webs scaling of each property varied little with latitude.
  This is consistent with the idea that species' niche breadths do not vary
  systematically with temperature and productivity but that the niche space
  might be larger in species-rich communities~\cite{Davies2007}. Rather than
  niche space depending on temperature and productivity, it may be that species
  diversity itself affects the biotic niche space available to species (although climate
  may affect speciation rates and therefore the diversity in a region~\cite{Currie2004}). 
  For example, as the plant diversity of a community increases both the 
  variety of food available to herbivores and the structural variety of the habitat will also increase.

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e., generality increased more steeply with
  increasing species richness) in higher-latitude food webs. In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and vulnerability and links per species. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower niches~\cite{Vazquez2004}
  and a higher proportion of specialists.

  That freshwater food webs supported the hypothesis of narrower niches in the
  tropics --while other ecosystem types did not-- is noteworthy given that
  these ecosystems (especially streams) are known for being highly variable
  and that seasonal variability is one of the proposed drivers of the
  latitude-niche breadth hypothesis~\cite{Vazquez2004}. Both streams and lakes
  can experience severe changes in water temperature and volume (e.g., floods,
  drying, freezing) that remove food or other resources (notably oxygen during
  freezing events)~\cite{Winterbourn1997,Meding2001}. These events are often
  linked to seasonal events such as snowmelts or summer
  drought~\cite{Winterbourn1997}). Further, both temperate streams and lakes
  tend to experience seasonal strong pulses of allochthonous inputs (e.g.,
  fallen leaves, terrestrial
  invertebrates~\cite{Nakano2001,Lennon2004,Zeng2008}. These trends combined
  mean that, relative to estuarine and marine communities, freshwater food 
  webs may experience high turnover in both community composition and 
  productivity~\cite{Tilzer1988,Magalhaes1993,Baird1989}. Notable exceptions from the above
  trends are New Zealand stream communities (representing 31 of the 71 stream
  food webs in our dataset), which experience unpredictable flooding and
  drying throughout the year and do not receive seasonally pulsed
  subsidies~\cite{Winterbourn1997,Winterbourn1981}. However, as this subset of
  webs is very tightly grouped in latitude ($44.64-46.41^{\circ}$S, within an
  overall range of $23.00-69.02^{\circ}$ for stream communities), it is unlikely that they have greatly
  influenced our results (see also \emph{Appendix S3}). Moreover, just as in
  highly-variable communities where said variation is more seasonal, New
  Zealand communities are dominated by ecological
  generalists~\cite{Winterbourn1997,Winterbourn1981} implying that they appear
  to fit the general pattern of streams worldwide.

  Importantly, while terrestrial communities are also strongly seasonal at
  high latitudes and can receive significant allochthonous
  inputs~\cite{Nakano2001}, terrestrial consumers tend to be
  morphologically specialised for feeding on particular prey~\cite{Liem1990}.
  This means that primarily gape-limited aquatic consumers tend to be more
  generalist across all types of aquatic habitats than terrestrial 
  consumers~\cite{Liem1990,Shurin2006}. This trend also held in our data, although the difference between generality was not significant (
  $\mu_{Aquatic}$=\st{5.47}5.86, $\mu_{Terrestrial}$=\st{3.82}3.75; $p$=\st{0.007}\textless0.001 for $\mu_{Aquatic}$\textgreater$\mu_{Terrestrial}$). 
  The key to this argument is therefore whether freshwater ecosystems experience
  more severe seasonal variation than marine and estuarine ecosystems. Although
  we are not aware of any study explicitly comparing seasonal variation in
  fresh and saltwater ecosystems at similar latitudes, we believe that freshwater 
  ecosystems are indeed likely to experience more severe changes because of their small size.
  While oceans and estuaries certainly vary in terms of water temperature and
  nutrients over the course of a year~\cite{Baird1989}, these changes are likely to be slower
  and milder than in freshwaters because marine and estuarine communities are buffered 
  by being open to the ocean rather than isolated in the midst of a terrestrial matrix. 
  Supporting this hypothesis, net primary productivity is much more variable over the course of a year 
  in non-marine communities~\cite{Field1998}. Thus suggests that niche breadths may also be 
  more variable over the course of the year.

\section*{Conclusion}

  Overall, our results were inconsistent with the latitude-niche breadth
  hypothesis in estuarine, marine, and terrestrial communities but consistent
  with the hypothesis of greater specialisation in the tropics in stream and
  lake food webs. This suggests that different mechanisms may structure food
  webs in different habitat types and that freshwater food webs in particular
  may be strongly affected by seasonal variation. Freshwater food webs also
  appear to have different predator-prey biomass ratios than other ecosystem
  types~\cite{Riede2011}; although it is not clear whether these ratios are related
  to seasonality, this could be a promising avenue for future research. In addition, different
  relationships between latitude and niche breadth in different habitat types
  go some way towards explaining the equivocal support for the opposing
  hypotheses of narrower niches in the tropics~\cite{Vazquez2004} and broader
  niche space in the tropics~\cite{Davies2007}. Our study indicates that both 
  have merit but would appear to apply to different systems.

\section*{Data Accessibility}

  Food webs used in this study were retrieved from the University of Canberra's 
  GlobalWeb database (\cite{GlobalWeb}; \emph{www.globalwebdb.com}) and from
  two papers (\cite{Riede2011}; \emph{http://dx.doi.org/10.1111/j.1461-0248.2010.01568.x}
  and \cite{Dunne2013}; \emph{http://dx.doi.org/10.1371/journal.pbio.1001579}). Original sources for the
  food webs are given in \emph{Appendix S1}.

\section*{Competing interests}

  We have no competing interests.

\section*{Authors' contributions}

  ARC, DBS, and TNR designed the study, ARC collected
  published data, performed the analyses, and wrote the
  first draft. DBS and TNR substantially revised the article;
  all authors approved the final version.

\section*{Acknowledgements}

  We thank members of the Romanuk, Stouffer, and Tylianakis labs for their comments on
  the manuscript. We also thank Angus McIntosh for valuable discussion on stream food webs and Matthias Schleuning, Robby Stoks, and one anonymous reviewer for their insightful questions.

\section*{Funding}
  This research was supported by an NSERC USRA undergraduate scholarship and NSERC PGS-D 
  graduate scholarship (to ARC), a Marsden Fund Fast-Start grant (UOC-1101) and a 
  Rutherford Discovery Fellowship, both administered by the Royal Society of New Zealand 
  (to DBS), and an NSERC Discovery Grant (to TNR).

\newpage
\bibliographystyle{prsb}
\bibliography{abbreviated}

@@

\section*{Figure Captions}

\begin{figure}[h]
\centerline{\includegraphics*[height=.5\textheight]{Figures/Other/conceptual_figure.eps}}
\caption{In A) we show the known scaling relationship between link density (links per species) and 
species richness. This scaling relationship is a power law and therefore linear in a log-log plot.
In B) we show two versions of the latitudinal-niche breadth hypothesis that have been proposed to
explain this gradient. Hypothesis 1 posits that greater environmental stability in the tropics will 
allow species to evolve narrower niches (indicated by parabolas) than those at the poles. Hypothesis 2 
suggests that species will have constant niche sizes over latitude but that greater primary productivity 
in the tropics creates a larger niche space such that each species still occupies a smaller proportion 
of the total niche space. These two hypotheses have different implications for the scaling of food-web 
properties such as the number of feeding links per species with species richness. C) If hypothesis 1 is 
true, then the exponent of the scaling relationship between link density and species richness should 
be larger towards the poles, where each additional species in the food-web will have a larger niche 
(i.e., more feeding links). If hypothesis 2 is true, then the exponent of this distribution should 
not vary significantly over latitude.}
\label{concept}
\end{figure}

\begin{figure}[h]
% \centerline{\includegraphics*[height=.5\textheight]{Figures/by_TL/scaling_with_S/proportions/S_fitlines_nonts_new.eps}}
\centerline{\includegraphics*[height=.5\textheight]{Figures/by_TL/scaling_with_S/proportions/STL_fitlines_nonts.eps}}
\caption{
Scaling relationships for re-scaled link density, generality, and vulnerability 
relative to the species richness of a food web. Link density, generality,
and vulnerability were each re-scaled to remove the effects of latitude and ecosystem
type. As these relationships take the form of power laws, we did this by dividing the food-web
property (e.g. link density) by species richness raised to an exponent including the 
effects of latitude and, where applicable, ecosystem type and the interaction between ecosystem
type and latitude. Note that in all cases estuarine food webs were treated as the baseline 
ecosystem type, but that at most two ecosystem types had interactions between ecosystem type and
latitude retained in the best-fit model (see \emph{Results} for specifics). For each relationship, 
we show the re-scaled values (white circles) as well as the overall scaling relationship using estuarine
ecosystems as a baseline (black line, N=196 food webs). For a figure with the uncorrected values,
see Fig.~\emph{S7}, \emph{Appendix S4}.}
\label{props_v_lat}
\end{figure}

\begin{figure}[h]
%\centerline{\includegraphics*[width=.85\textwidth]{Figures/by_TL/marginal/S_marginal_latitude_proportions.eps}}
% \centerline{\includegraphics*[width=.85\textwidth]{Figures/by_TL/marginal/S_marginal_latitude_non_TS_updated.eps}}
\centerline{\includegraphics*[width=.85\textwidth]{Figures/by_TL/marginal/S_marginal_latitude_non_TS_corrected.eps}}
\caption{Changes to the scaling of link density, generality, and vulnerability with species richness across ecosystem
types and over latitude. We show the estimated scaling exponent for species richness (black
line) with its 95\% confidence interval (in grey), based on N=196 empirical food webs.
Latitude is given in degrees from the equator regardless of direction.} \label{S} \end{figure}

>>>>

% \linenumbers
\begin{spacing}{2.0}

\section*{Correction}

  A subset of the food webs (75/263 webs) used in these analyses contained formatting errors or irregularities that meant they were not properly included in the original analyses. We have corrected these webs (see \emph{Table S2, Appendix S1} for details) and repeated all analyses. The corrected results were very similar to those in the original manuscript: original values (struck-through) and corrected values are provided below.

\subsection*{Results}

  Link density (mean number of feeding links per species), generality (mean
  number of prey per species), and vulnerability (mean number of predators per
  species) were strongly and positively correlated ($R^2$=\st{0.891}0.915 for link
  density and generality, $R^2$\textgreater0.999 for link density and
  vulnerability, and $R^2$=\st{0.890}0.912 for generality and vulnerability). Contrary
  to the expected latitudinal gradient, the best-fit version of
  equation (1) did not include a significant
  effect of latitude on species richness for any ecosystem type. Further, there were no significant
  relationships between link density, generality, or vulnerability with
  latitude for any ecosystem type.

  Each measure of specialisation increased with increasing
  species richness \\($\beta_0$=\st{0.637}0.688, $p$\textless0.001; $\beta_0$=\st{0.553}0.658,
  $p$\textless0.001; and $\beta_0$=\st{0.637}0.695, $p$\textless0.001, respectively;
  Fig.~\ref{props_v_lat}). For estuarine, marine, and terrestrial food webs the
  strength of this scaling did not vary with latitude
  ($\beta_{Latitude}$=-0.001, $p$=\st{0.365}0.210 for link density;
  $\beta_{Latitude}$=-0.001, $p$=\st{0.535}0.315 for generality; and
  $\beta_{Latitude}$=-0.001, $p$=\st{0.366}0.205 for vulnerability; Fig.~\ref{S}). In
  lake food webs, however, the scaling of each property was stronger towards
  the poles ($\beta_{Latitude:Lake}$=\st{0.004, $p$=0.019}0.003, $p$=0.031;
  $\beta_{Latitude:Lake}$=0.005, $p$=\st{0.004}0.005; and
  $\beta_{Latitude:Lake}$=\st{0.004, $p$=0.018}0.004, $p$=0.030, respectively). In stream food
  webs, generality increased more rapidly towards the poles
  ($\beta_{Latitude:Stream}$=\st{0.007, $p$=0.001}0.005, $p$=0.008) while link density and
  vulnerability did not vary with latitude (i.e., the interaction term 
  $\beta_{Latitude:Stream}$ was not retained in the best-fit models).

\subsection*{Discussion}

  ... primarily gape-limited aquatic consumers tend to be more
  generalist across all types of aquatic habitats than terrestrial 
  consumers~\cite{Liem1990,Shurin2006}. This trend also held in our data, although the difference between generality was not significant (
  $\mu_{Aquatic}$=\st{5.47}5.86, \\$\mu_{Terrestrial}$=\st{3.82}3.75; $p$=\st{0.007}\textless0.001 for $\mu_{Aquatic}$\textgreater$\mu_{Terrestrial}$).

@@

\bibliographystyle{prsb}
\bibliography{abbreviated}

@@

\section*{Figures}

\begin{figure}[h]
% \centerline{\includegraphics*[width=\textwidth]{Figures/by_TL/scaling_with_S/proportions/STL_fitlines_nonts.eps}}
% \centerline{\includegraphics*[width=\textwidth]{Figure2_updated.eps}}
\caption{
Scaling relationships for re-scaled link density, generality, and vulnerability 
relative to the species richness of a food web. Link density, generality,
and vulnerability were each re-scaled to remove the effects of latitude and ecosystem
type. As these relationships take the form of power laws, we did this by dividing the food-web
property (e.g. link density) by species richness raised to an exponent including the 
effects of latitude and, where applicable, ecosystem type and the interaction between ecosystem
type and latitude. Note that in all cases estuarine food webs were treated as the baseline 
ecosystem type, but that at most two ecosystem types had interactions between ecosystem type and
latitude retained in the best-fit model (see \emph{Results} for specifics). For each relationship, 
we show the re-scaled values (white circles) as well as the overall scaling relationship using estuarine
ecosystems as a baseline (black line, N=196 food webs). For a figure with the uncorrected values,
see Fig.~\emph{S7}, \emph{Appendix S4}.}
\label{props_v_lat}
\end{figure}

\begin{figure}[h]
% \centerline{\includegraphics*[width=\textwidth]{Figures/by_TL/marginal/S_marginal_latitude_non_TS_corrected.eps}}
% \centerline{\includegraphics*[width=\textwidth]{Figure3_updated.eps}}
\caption{Changes to the scaling of link density, generality, and vulnerability with species richness across ecosystem
types and over latitude. We show the estimated scaling exponent for species richness (black
line) with its 95\% confidence interval (in grey), based on N=196 empirical food webs.
Latitude is given in degrees from the equator regardless of direction.} \label{S} \end{figure}

### 2018-08-08 manuscript/erratum_letter.tex
uploading a shortened correction

I am writing regarding my article \emph{Specialisation in food webs scales with species richness but not with latitude} (article ID \#RSPB-2015-1589.R2, doi: 10.1098/rspb.2015.1589, published November, 2015). Recently I have revisited the dataset used in this article and noticed that 75  of the webs we used (out of a sample of 263) have irregularities or errors of formatting that caused them to be incorrectly processed by our original code. Fortunately, after repeating all analyses with corrected versions of these 75 webs our results were qualitatively identical to those published and our conclusions remain unchanged. Unfortunately, the specific values given in the published text are nearly all incorrect.

I have attached an updated version of the manuscript with the original values struck through and the corrected values provided side-by-side. I hope that this demonstrates that our conclusions hold and will be suitable as a corrigendum for our manuscript. I have also attached an updated supplemental information document which now includes a table (Table S2) giving details of the webs which required corrections.

>>>>

I am writing regarding my article \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types.} (article ID \#RSPB-2015-1589.R2). The full citation details for this article are: Cirtwill, A.R., Stouffer, D.B., \& Romanuk, T.N. (2015) \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types}. \textbf{Proceedings of the Royal Society B}: 282(1819), 20151589. doi: 10.1098/rspb.2015.1589

Recently I have revisited the dataset used in this article and noticed that 75  of the webs we used (out of a sample of 263) have irregularities or errors of formatting that caused them to be incorrectly processed by our original code. Fortunately, after repeating all analyses with corrected versions of these 75 webs our results were qualitatively identical to those published and our conclusions remain unchanged. Unfortunately, the specific values given in the published text are nearly all incorrect.

I have attached a corrected version of the results and a small portion of the discussion with the original values struck through and the corrected values provided side-by-side. No further changes to the manuscript were necessary. I hope that this demonstrates that our conclusions hold and will be suitable as a corrigendum for our manuscript. I have also attached an updated supplemental information document which now includes a table (Table S2) giving details of the webs which required corrections.
